### 2021-02-15 Modules/sec_future_wavfrom.tex
rev03

In practical RADAR related systems, the waveform of choice is pulse (chirp) based.
It holds some obvious advantages over the CW transmission such as
\begin{itemize}
    \item Lower power consumption
    \item Possible match filters should allow the decision of weather to re-transmit or not
    \item In multi target scenario, there is a smaller chance of interference.
    \item The most interesting research option is the use of several (possibly many) independent harmonics as super positioned multiple dual frequency beamformers, which simultaneously scan a wide area with multiple resolutions, by merely selecting harmonic couples and filter coefficients, where each harmonic/coefficient set it pointed to a specific area.
    In the expense of computational effort (without adding elements to the array), it seems possible that one may have simultaneous scan of the entire arena in any wanted direction/range/resolution by mere choosing multiple frequency couples which are simultaneously processed independently.
    \par 
    As a simple example, consider that the arena consists of $M$ zones of interest, where each zone may be described as a angular bounded radial slice (similar to the visualization of Fig.~\ref{fig_rangeAzimuthSelectivity}) which may be described as $$a_{m}=\cBrace{\vecnot{p}\ \middle\vert\  \theta_{d,m,\text{min}}<\text{DOA}\rBrace{\vecnot{p}}<\theta_{d,m,\text{max}}\ ,\ R_{m,\text{min}}<\norm{\vecnot{p}}<R_{m,\text{max}}}\ \forall\ m=0\dots{}M-1.$$
    \par
    Each zone will be assigned with matching couple of carrier frequencies ($\omega_{1,2}$) to cover its radial width (where the radial width is equivalent to $R_{\text{rt}}$ of \eqref{eq_DF_phErr}) and coefficient sets ($\vecnot{\alpha}_{m,i},\vecnot{\beta}_{m,i} \ \forall\ m=0\dots{}M-1, i=1,2$), which cover the zone's angular diameter which is equivalent to determining the desired beamwidth via setting the proper $\kappa$ of \eqref{eqn_H_DF_CB}.
    \par
    Leaving aside the needed computational effort and physical limitations of transmitting multiple narrowband signals, assuming large $M$ values, we have a super-positioned feedback beamformer which simultaneously scans the entire arena.
\end{itemize}

>>>>

In practical RADAR related systems, the waveform of choice is pulse (in most cases, chirp) based.
It holds some obvious advantages over the CW transmission such as
\begin{itemize}
    \item Lower power consumption due to lower transmissions duty cycle. 
    \item In multi target scenario, there is a smaller chance of interference for the reflected signals from the targets temporally distinct if the targets are in different distances.
    \item The use of wide band signals may as in RADAR systems, increase spatial resolution in the radial axis.
\end{itemize}
Another interesting option is the using several (possibly many) independent harmonics as super positioned multiple dual frequency beamformers, which simultaneously scan a wide area with multiple resolutions.
By merely selecting harmonic couples and filter coefficients, each harmonics pair is pointed to a specific area.
It follows that with the expense of computational effort (without adding elements to the array), it seems possible that one may have simultaneous scan of the entire arena in any wanted direction/range/resolution.
\par 
As a simple example, consider that the arena consists of $M$ zones of interest, where each zone may be described as a angular bounded radial slice (similar to the visualization of Fig.~\ref{fig_rangeAzimuthSelectivity}) which may be described as $$a_{m}=\cBrace{\vecnot{p}\ \middle\vert\  \theta_{d,m,\text{min}}<\text{DOA}\rBrace{\vecnot{p}}<\theta_{d,m,\text{max}}\ ,\ R_{m,\text{min}}<\norm{\vecnot{p}}<R_{m,\text{max}}}\ \forall\ m=0\dots{}M-1.$$
\par
Each zone will be assigned with matching couple of carrier frequencies ($\omega_{1,2}$) to cover its radial width (where the radial width is equivalent to $R_{\text{rt}}$ of \eqref{eq_DF_phErr}) and coefficient sets ($\vecnot{\alpha}_{m,i},\vecnot{\beta}_{m,i} \ \forall\ m=0\dots{}M-1, i=1,2$), which cover the zone's angular diameter which is equivalent to determining the desired beamwidth via setting the proper $\kappa$ of \eqref{eqn_H_DF_CB}.
\par
Leaving aside the needed computational effort and physical limitations of transmitting multiple narrowband signals, assuming large $M$ values, we have a super-positioned feedback beamformer which simultaneously scans the entire arena.

### 2021-02-15 Modules/sec_prlm_FIR_IIR.tex
rev03

The design of the digital filter is very well established and thoroughly studied research field.
As stated in \cite{oppenheim1975digital} (and numerous other literature), two key filter types are used - i.e. Finite Impulse Response (FIR) and Infinite Impulse Response (IIR) filters.

@@

\par Each digital filters belong in one of two architectures

@@

\par It is well known \cite{rabiner1974some} that IIR filters have several appealing advantages over their matching (in number of elements) FIR versions.
A well known \cite{oppenheim1975digital} fact in digital signal processing is that the filter latency may be frequency dependant, namely linearly dependent in the phase response derivative of the specific frequency.

@@

Due to this important property, many applications which are sensitive to signal distortions, especially communication based systems, chose the FIR architecture.

@@

\par The main advantage of IIR filters is that for a given set of frequency response specifications, the IIR filter will (in most cases) require substantially less coefficients, which may even reach few order of magnitude \cite{rabiner1974some}.
Concluding that spatial linear phase response in DOA domain is not a necessity, combined with the known efficiency of the IIR architecture and the need for smaller and cheaper arrays, the motivation for finding the spatial version of the temporal IIR is complete.

@@

As opposed to the standard scheme, we suggest a continuous re-transmission of the signal and its echoes back to the platform, generating a spatial feedback loop between the array and the target.

>>>>

The design of the digital filters is very well established and thoroughly studied research field.
As stated in \cite{oppenheim1975digital} (and numerous other literature), two key filter types are Finite Impulse Response (FIR) and Infinite Impulse Response (IIR) filters.

@@

@@

\par It is well known \cite{rabiner1974some} that IIR filters have several appealing advantages over their matching (in terms of number of taps) FIR counterparts.
The main advantage of IIR filters is that for a given set of frequency response specifications, the IIR filter will (in most cases) require substantially less coefficients, which may even reach few order of magnitude \cite{rabiner1974some}.
In the spatial processing context, this implies improved spatial performance, using the same or smaller number of sensors.
However, a well known \cite{oppenheim1975digital} fact in digital signal processing is that the filter latency may be frequency dependant, namely linearly dependent in the phase response derivative of the specific frequency.

@@

Due to this important property, many applications which are sensitive to signal distortions, especially communication based systems, chose the FIR architecture, for it is possible to design it as linear phase filter.

@@

% \par
% Concluding that spatial linear phase response in DOA domain is not a necessity, combined with the known efficiency of the IIR architecture and the need for smaller and cheaper arrays, the motivation for finding the spatial version of the temporal IIR is complete.

@@

As opposed to the standard scheme, we suggest in Chapter.~\ref{chap:firstchap} a continuous re-transmission of the signal and its echoes back to the platform, generating a spatial feedback loop between the array and the target.

### 2021-02-15 Modules/sec_prlm_array_perf.tex
rev03

In this work, we focus on localization related applications and suggest a new architecture.
Hence, to analyse the presented scheme, we use classic localization-related performance measurements as the comparison is to known localization related array processing schemes.

@@

  In Fig.~\ref{fig_common_bps1}, the response is presented on the interval of 0 to $\pi$ with units of dB.

@@

Although may be dependant in various parameters, for the sake of simplicity, we consider a planar (i.e. azimuth without elevation) DOA-only dependant array response (i.e. the \emph{beampattern})

@@

It's characteristics, such as lobes width, attenuation and positioning are then analysed and compared between different schemes.
The analysis can theoretical, when the beampattern's mathematical expression is known or graphical analysis for empirical scenarios.
In this work, as we deal with 2D planar localization problems, we find the presentation of Fig.~\ref{fig_common_bps1} most suitable.
The horizontal axis, being the DOAs of the impinging signal, may be presented in various ways where units may be degrees radians etc.
Also, as the DOA is periodical one should also choose the centering, for example when the beampattern is steered towards different angles and the author wants to center the mainlobe.
In this work, we follow choose to use \cite{van2004optimum}'s $\psi$-space, i.e.

@@

The reader may notice that $\psi$ is exactly the electric phase $\theta$ defined in \eqref{eq:thetaULA}. 
Also, we choose to plot beampatterns using $\Delta\psi = \psi\rBrace{\theta_{d}} - \psi\rBrace{\theta_{s}}$, where $\theta_{s}$ is the steering angle of the array, for it centers the mainlobe, enabling convenient comparison.

@@

As the absolute output gain is system dependent, comparing two different is not possible where their maximal gain of the mainlobe are different.

@@

This quantity will be referred as $\mathcal{H}$.

@@

Considering localization problems, possibly the main interest is spatial resolution, related to localization accuracy and distinction of spatially close objects, where ``close`` can be interpreted as small physical distance between objects of interest or objects that may be physically distant from each other but from the array's perspective are in relatively similar.
Relating to the beampttern, the spatial resolution is obviously linked to the main lobe's width, i.e. higher resolution is acheived as the main lobe thins. 
Hence to unify the analysis, the Half Power BeamWidth (HPBW) is defined to be the 3-dB beamwidth, i.e. the point where $\abs{B\rBrace{\psi}}^{2} = 0.5$ or $\abs{B\rBrace{\psi}} = 1/\sqrt{2}$.

@@

As the beampattern consists of multiple harmonics (as the number of elements $N$), sidelobes occur when the some harmonics are summed, showing as peaks other than the mainlobe.
As this is an unwanted result, the height and their rate of decrease are measured aiming to prefer arrays where their height is reduced and the decrease rate is increased.

@@

The array directivity $\mathcal{D}$ \cite{van2004optimum}, defined as

@@

    \mathcal{D}\rBrace{N,r} = \frac{\Hr_{\Delta\theta=0,\Delta\phi=0,r}}{\frac{1}{2\pi}\int_{0}^{2\pi}\Hr_{\Delta\theta,\Delta\phi=0,r}\ d\Delta\theta} = \frac{2\pi}{\int_{0}^{2\pi}\Hr_{\Delta\theta,\Delta\phi=0,r}\ d\Delta\theta},

@@

measures the ratio between the maximal array gain at its mainlobe, to the average gain over all directions. 
For uniformly weighted ULAs with no feedback, it is known \cite{van2004optimum} that $\mathcal{D}\rBrace{N,0} = N$.

>>>>

In this work, we focus on localization related applications and suggest a new architecture for the beamformer design.
Hence, to analyse the presented scheme, we use classic localization-related performance metrics.

@@

  In Fig.~\ref{fig_common_bps1}, the response is presented on the $\vBrace{0,\pi}$ interval with units of dB.

@@

Although for sensors in 3D space the DOA may depend on two angles (azimuth and elevation), for the sake of simplicity, we consider a planar (i.e. azimuth without elevation) DOA-only dependant array response (i.e. the \emph{beampattern})

@@

It's characteristics, such as main lobe width, position and attenuation of sidelobes are then analysed and compared between different schemes.
The analysis can be done theoretically, when the beampattern's mathematical expression is known or it can be conducted numerically for empirical scenarios.
% In this work, as we deal with 2D planar localization problems, we find the presentation of Fig.~\ref{fig_common_bps1} most suitable.
The DOAs of the impinging signal, may be presented in various ways where units may be degrees or radians.
% Also, as the DOA is periodical one should also choose the centering, for example when the beampattern is steered towards different angles and the author wants to center the mainlobe.
In this work, we choose to follow \cite{van2004optimum}'s $\psi$-space, i.e.

@@

The reader may notice that $\psi$ is exactly the electric phase $\theta$ defined in \eqref{eq:thetaULA} therefore, throughout the rest of the work, we use the $\theta$ notation. 
Also, we choose to plot beampatterns using $\Delta\theta = \theta\rBrace{\theta_{d}} - \theta\rBrace{\theta_{s}}$, where $\theta_{s}$ is the steering angle of the array, for it centers the mainlobe, enabling convenient comparison.

@@

As the absolute output gain is system dependent, comparing two different beampatterns is not possible where their maximal gain of the mainlobe are different.

@@

This quantity will be referred to as $\mathcal{H}$.

@@

% Considering localization problems, possibly the main interest is spatial resolution, related to localization accuracy and distinction of spatially close objects, where ``close`` can be interpreted as small physical distance between objects of interest or objects that may be physically distant from each other but from the array's perspective are in relatively similar.
% \par
Considering localization problems, the spatial resolution is obviously linked to the main lobe's width, i.e. higher resolution is achieved for narrower main lobe.
The Half Power BeamWidth (HPBW), marked as $\Delta\theta_{HPBW}$, is defined to be the 3-dB beamwidth, i.e. the point where $\abs{B\rBrace{\theta}}^{2} = 0.5$ or $\abs{B\rBrace{\theta}} = 1/\sqrt{2}$.

@@

% As the beampattern consists of multiple harmonics (as the number of elements $N$), sidelobes occur when the some harmonics are summed, showing as peaks other than the mainlobe.
% As this is an unwanted result, the height and their rate of decrease are measured aiming to prefer arrays where their height is reduced and the decrease rate is increased.
A known \cite{van2004optimum} parasitic phenomenon of beampatterns is sidelobes, which manifests as energy peaks outside the mainlobe, typically highest near the mainlobe and decreasing towards the deges of the beampattern.
In the context of localization, this can cause false detection or reduction in spatial resolution.
Related spatial performance metrics are the sidelobes gain and rate of decrease

@@

A common measure of performance of an array or aperture is the directivity $\mathcal{D}$.
As in \cite{van2004optimum}, in a transmitting array or aperture, $\mathcal{D}$ represents the maximum radiation intensity (power per DOA) divided by the average radiation intensity (averaged over all DOAs).
Also, in a receiving antenna, the denominator represents the noise power at the array (or aperture) output due to isotropic noise (noise distributed uniformly over all DOAs). 
The numerator will represent the power due to a signal arriving from a certain DOA ($\theta_{T}$) - Thus, $\mathcal{D}$ can be interpreted as the array gain against isotropic noise.
We define the power pattern, $P\rBrace{\theta_{d}}$, to be the squared magnitude of the beampattern $B\rBrace{\theta_{d}}$
    P\rBrace{\theta_{d}} = \abs{B\rBrace{\theta_{d}}}^{2}
where the frequency dependence is suppressed.
Then the directivity $\mathcal{D}$ is defined as

@@

    \mathcal{D}\rBrace{\theta_{T}} = \frac{
    P\rBrace{\theta_{T}}
    }{
    \frac{1}{2\pi}\int_{0}^{2\pi}P\rBrace{\theta}d\theta
    }

@@

In the following, as we consider steered (to $\theta_{s}$) arrays, we suppress the DOA dependency such that $\mathcal{D} = \mathcal{D}\rBrace{{\theta_{s}}}$. 
For uniformly weighted ULAs with no feedback, it is known \cite{van2004optimum} that $\mathcal{D} = N$.

### 2021-02-15 Modules/sec_prlm_localization.tex
rev03

Following the classical methods overview in \cite{krim1996two},localization estimators may be divided to two main groups - i.e. \textit{spectral-based} and \textit{parametric}.

@@

This approach of spatial filtering, is considered to suffers from fundamental

@@

Although trials were made to increase the resolution of an array given a certain aperture, this notion remains intact to these days.
In this contribution we tackle this statement as will shown in Chapter\ref{chap:firstchap}.
The later approach, of parameter estimation assumes a physical model of the experimental scenario, which serve a basis for the development of spatio-temporal estimators which takes advantage not only from the spatial diversity but also from the temporal information collected via sequential time separated snapshots in order to increase the overall SNR.
Such methods, not only enabled the increase of spatial filtering resolution, but also allowed the implementation of data processing algorithms, more sophisticated than a mere search with a steered array.

@@

The earliest localization algorithms, which date back to world war II, are the beamforming based, e.g. the conventional Bartlet \cite{van2004optimum} and Capon's \cite{capon1969high} beamformers.
Early examples of parametric estimators, are the Maximum-Likelihood \cite{macdonald1969optimum,schweppe1968sensor} and Maximum-Entropy \cite{ables1974maximum} which assume a known PDF to the received signal and estimate the desired parameters according to empirical results and fitting the closest matching PDF.

@@

MUSIC essentially relieved the designer from the need of accurate radiation patterns by exploiting the reduction in analytical complexity that could be achieved by calibrating the array.
Thus, the array response calculation task was reduced to that of measuring and storing the response. 

@@

Following state-of-the-art developments \cite{BOOK-classicalAndModernDOA}, the sub-space based algorithms are still modified and adjusted to specific scenarios as in \cite{LpNorm MUSIC} but also new and interesting approaches emerged

@@

    \item High Order Statistics (HOS, also refereed as cumulants), which extracts more information from the samples' higher order moments, is a very active field of research due to some fundamental issues which are inherently resolved - i.e. the Gaussian noise vanishes in the 4th order statistics and the ability to resolve more DOAs than array elements \cite{chevalier2006high}.
    This approach, being costly in computation effort, became popular probably due to the recently available low-cost powerful computation.
    \item Inspired by the vast research field of sparse representations some algorithms \cite{nadiri2014localization} use non-Frobenius norm (i.e. $L^{p}$ norms where $p<2$), in order improve estimation resolution under high reverberation acoustic scenarios.

@@

    \item Naturally, also many trials of harnessing the promising concept of neural networks are being done, for example \cite{shareef2008localization}.

>>>>

Following the classical methods overview in \cite{krim1996two}, localization estimators may be divided to two main groups - i.e. \textit{spectral-based} and \textit{parametric}.

@@

This approach of spatial filtering, is considered to suffer from fundamental

@@

Although trials were made to increase the resolution of an array given a certain aperture, this limitation remained intact.
In this contribution we tackle this issue as will shown in Chapter\ref{chap:firstchap} by introducing a spatial feedback.
The later approach, of parameter estimation assumes a physical model of the experimental scenario, which serve a basis for the development of spatio-temporal estimators which take advantage not only from the spatial diversity but also from the temporal information collected via sequential time separated snapshots in order to increase the overall SNR.
Such methods, not only enabled the increase of spatial filtering resolution, but also allowed the implementation of data processing algorithms, which are more sophisticated than a mere search with a steered array.

@@

The earliest localization algorithms, which date back to world war II, are beamforming based, e.g. the conventional Bartlet \cite{van2004optimum} and Capon's \cite{capon1969high} beamformers.
Early examples of parametric estimators, are the Maximum-Likelihood \cite{macdonald1969optimum,schweppe1968sensor} and Maximum-Entropy \cite{ables1974maximum} which assume a known PDF to the received signal and estimate the desired parameters according to sampled data and fitting the closest matching PDF.

@@

MUSIC essentially relieved the designer from designing accurate radiation patterns by introducing the concept of array calibration.

@@

Following state-of-the-art developments \cite{tuncer2009classical}, the sub-space based algorithms are still modified and adjusted to specific scenarios as in \cite{LpNorm_MUSIC} but also new and interesting approaches emerged such as

@@

    \item High Order Statistics (HOS), which extracts more information from the samples' higher order moments, is a very active field of research due to some fundamental issues which are inherently resolved - i.e. the Gaussian noise vanishes in the 4th order statistics and the ability to resolve more DOAs than array elements \cite{chevalier2006high}.
    This approach, being costly in computation effort, became popular probably due to the recently available low-cost powerful computation platforms.
    \item Inspired by the vast research field of sparse representations, some algorithms \cite{nadiri2014localization} use $L^{p}$ norms (where $p<2$), in order to improve estimation resolution under high reverberation acoustic scenarios.

@@

    \item Naturally, also many trials of harnessing the promising concept of neural networks are being done, see for example \cite{shareef2008localization}.

### 2021-02-15 Modules/sec_prlm_propWaveField.tex
rev03

An elementary physical phenomenon is the spatial dynamics of waves, referred to as wave-propagation.

@@

@@

Denoting $t$ as the time, the time-space representation of the a signal is $f\rBrace{t,\vecnot{r}}$ or , where the relations between the Cartesian and spherical coordinates are given in Fig.~\ref{fig_coordinates}.

@@

A possible solution to \eqref{eq_prlm_waveEq}, $f_p\rBrace{t,x,y,z}$, may be of a complex exponential form:

@@

f_p\rBrace{t,x,y,z} = A\exp{j\rBrace{\omega{t}-k_{x}x-k_{y}y-k_{z}z}}

@@

Where $A$ is a complex constant, $\omega$ denotes temporal radial frequency and $k_{x}x,k_{y},k_{z}$ are real constants. 
Plugging \eqref{eq_prlm_waveEq_pSol} into \eqref{eq_prlm_waveEq} results in the \emph{monochromatic plane wave}:

@@

k_{x}^{2}-k_{y}^{2}-k_{z}^{2} = \frac{\omega^{2}}{c^{2}}.

@@

Using the plane wave notation emphasizes the fact that for a given time, $t_{0}$, all points on a plane given by $k_{x}x+k_{y}y+k_{z}z = constant$ are with the same wave value and the vector notation of \eqref{eq_prlm_waveEq}'s solution is

@@

f\rBrace{t,\vecnot{x}}=A\exp{j\rBrace{\omega{t}-\vecnot{k}\vecnot{x}}}.

@@

where $\vecnot{k}\vecnot{x}=constant$ are planes of constant wave value.

@@

\omega\delta{t}-\vecnot{k}\delta\vecnot{x}=0.

@@

Assuming $\delta\vecnot{x}$ and $\vecnot{k}$ have the same direction, $\vecnot{k}\delta\vecnot{x} = \abs{\vecnot{k}}\abs{\delta\vecnot{x}}$ and $\frac{\delta{\vecnot{x}}}{\delta{t}} = \frac{\omega}{\abs{\vecnot{k}}}$, where $\frac{\delta{\vecnot{x}}}{\delta{t}}$ can designate the propagation speed of the plane wave. 

@@

\frac{\delta{\vecnot{x}}}{\delta{t}}=c,

@@

where $c>0$ is the wave velocity in the medium.
The \emph{wavelength} ($\lambda$) denote the distance the plane wave propagated during a single temporal period of $T=\frac{2\pi}{\omega}$.
Let $\vecnot{k}$ denote the \emph{wavelength vector}. 

@@

\lambda=\delta{\vecnot{x}}=\frac{2\pi}{\abs{\vecnot{k}}}.

@@

The spatial diversity of the ULA elements is expressed as a TOA difference of the impinging signal in each sensor which is will be shown to provide clues for the signal's DOA as can be seen in Fig.~\ref{fig_ULA_sketch}.

@@

When also considering the narrowband scenario, where TOA difference is merely phase shift, it seems obvious that by phase measuring, one should be able to extract the DOA from the spatially sampled data.

>>>>

An elementary physical phenomenon is the spatial and temporal dynamics of waves, referred to as wave-propagation.

@@

The relations between the coordinates are given in Fig.~\ref{fig_coordinates}.

@@

Denoting $t$ as the time, the time-space representation of the a signal is $f\rBrace{t,\vecnot{r}}$.

@@

A possible solution to \eqref{eq_prlm_waveEq}, $f_p\rBrace{t,x,y,z}$, is a complex exponential of the form:

@@

f_p\rBrace{t,x,y,z} = A\exp\rBrace{j\rBrace{\omega{t}-k_{x}x-k_{y}y-k_{z}z}}

@@

Where $A$ is a complex constant, $\omega$ denotes temporal radial frequency, $\vecnot{k} = \vBrace{k_{x},k_{y},k_{z}}^{T}$ denote the \emph{wavelength vector} and $k_{x},k_{y},k_{z}$ are real constants. 
Indeed, plugging \eqref{eq_prlm_waveEq_pSol} into \eqref{eq_prlm_waveEq} results in the \emph{monochromatic plane wave} satisfying:

@@

k_{x}^{2}+k_{y}^{2}+k_{z}^{2} = \frac{\omega^{2}}{c^{2}}.

@@

Using the plane wave notation 

@@

f\rBrace{t,\vecnot{x}}=A\exp{j\rBrace{\omega{t}-\vecnot{k}^{T}\vecnot{x}}},

@@

emphasizes the fact that for a given time, $t_{0}$, all points on a plane given by $k_{x}x+k_{y}y+k_{z}z = constant$ are with the same wave value where $\vecnot{k}\vecnot{x}=constant$ are planes of constant wave value.

@@

\omega\delta{t}-\vecnot{k}^{T}\delta\vecnot{x}=0.

@@

Assuming $\delta\vecnot{x}$ and $\vecnot{k}$ have the same direction, $\vecnot{k}^{T}\delta\vecnot{x} = \abs{\vecnot{k}}\abs{\delta\vecnot{x}}$ and $\frac{\abs{\delta{\vecnot{x}}}}{\delta{t}} = \frac{\omega}{\abs{\vecnot{k}}}$, where $\frac{\abs{\delta{\vecnot{x}}}}{\delta{t}}$ can designate the propagation speed of the plane wave. 

@@

\frac{\abs{\delta{\vecnot{x}}}}{\delta{t}}=c.

@@

The \emph{wavelength} ($\lambda$) denotes the distance the plane wave propagates during a single temporal period of $T=\frac{2\pi}{\omega}$.

@@

\lambda=\abs{\delta{\vecnot{x}}}=\frac{2\pi}{\abs{\vecnot{k}}}.

@@

The spatial diversity of the ULA elements is expressed as a TOA difference of the impinging signal in each sensor which will be shown to provide clues for the signal's DOA as can be seen in Fig.~\ref{fig_ULA_sketch}.

@@

When also considering the narrowband scenario, where TOA difference is merely phase shift, it seems obvious that by measuring phases, one should be able to extract the DOA from the spatially sampled data.

### 2021-02-15 Modules/sec_prlm_sensorArrays.tex
rev03

Sensor arrays are sets of independently positioned sensors, where each sensor is sampling temporal snapshots of impinging signals.
The samples from the entire array are then fused together in order to extract the underlying data.
The spatial diversity of the sampled data allows the extraction of spatial features which are not accessible by mere temporal analysis of the data.
Sensor arrays are used in numerous applications, ranging from source localization, through communication, medical applications, astronomy etc.
\par 
When designing a sensor array, multiple consideration are to be taken into account, ranging from physical characteristics e.g. area, weight and carrying platform, through performance related considerations such as accuracy, spatial selectivity, signal to noise ratio and even the financial aspect.
Obviously, the most fundamental array specification is its geometry, for it dictates the spatial relations between simultaneous samples measured by the array's sensors, which also greatly influences the processing methods of the raw-data.
A basic example of sensor array is the ULA (see Fig.~\ref{fig_ULA}), where its elements are uniformly spaced on a straight line with distance $d$ between a sensor to its nearest neighbour.
Considering an array of $N$ elements, the sensors are positioned at
\begin{equation}
p_{n}=nd,\ n=0,\dots,N-1
\end{equation}
where $n$ is the sensors index; $p_{n}$ denotes the position of the $n$'th sensor,$p_{0} = 0$ corresponds to the left-hand side of the array.
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/arrayBasic.png}
        \put (4, 4.5) {\tiny{$N$}}
        \put (86, 4.5) {\tiny{$N-1$}}
        \put (83.25, 28.5) {\tiny{$\rBrace{N-1}\cdot{}d$}}
        \end{overpic}
    \end{center}
     \caption{A uniform linear array of size $N$ and spacing $d$.}
    \label{fig_ULA}
\end{figure}
\par To proper formulate the discussed additional spatial information when using sensor array based processors, a quick overview on wave propagation is due and presented in the following section (Sec.~\ref{sec:prlm_propWaveField}).

>>>>

Sensor arrays are sets of independently positioned sensors, where each sensor is sampling temporal snapshots of impinging signals.
The samples from the entire array are then fused together in order to extract the underlying data.
The spatial diversity of the sampled data allows the extraction of spatial features.
Sensor arrays are used in numerous applications, ranging from source localization, communication, medical applications, astronomy etc.
\par 
When designing a sensor array, multiple consideration are to be taken into account, ranging from physical characteristics e.g. area, weight and carrying platform, through performance related considerations such as accuracy, spatial selectivity, signal to noise ratio and even the financial aspect.
The most fundamental array specification is its geometry, for it dictates the spatial relations between simultaneous samples measured by the array's sensors, which also greatly influences the processing methods of the raw-data.
A basic example of a sensor array is the ULA (see Fig.~\ref{fig_ULA}), where its elements are uniformly spaced on a straight line with distance $d$ between each pair of sensors.
Considering an array of $N$ elements, the sensors are positioned at
\begin{equation}
p_{n}=nd,\ n=0,\dots,N-1
\end{equation}
where $n$ is the sensors index; $p_{n}$ denotes the position of the $n$'th sensor, $p_{0} = 0$ corresponds to the left-hand side of the array as illustrated in Fig.~\ref{fig_ULA}.
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width=0.5\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/arrayBasic.png}
        \put (4, 4.5) {\tiny{$N$}}
        \put (86, 4.5) {\tiny{$N-1$}}
        \put (83.25, 28.5) {\tiny{$\rBrace{N-1}\cdot{}d$}}
        \end{overpic}
    \end{center}
     \caption{A uniform linear array of size $N$ and spacing $d$.}
    \label{fig_ULA}
\end{figure}
\par To properly formulate the discussed additional spatial information when using sensor arrays, a quick overview on wave propagation is due and presented in the following section (Sec.~\ref{sec:prlm_propWaveField}).
